\LARGE \bf Planning with Attitude

### 2020-10-14 paper/main.tex
tiny edits

    The Jacobian of $r(q)$ can be found using \eqref{eq:quat_jacobian} and \eqref{eq:quaternion_rotation}: \todo{Kevin: fix the second line of this equation:}
    \begin{equation} \label{eq:newton_foc}
        \begin{aligned}
            \nabla r(q) &= \pdv{r}{q} G(q) \\
              &= \left( \sum_i -2 H^T R(q)^T R({}^B \hat{v}_i) \right).
        \end{aligned}
    \end{equation}
    Given a guess solution, $q_k$, The standard Gauss-Newton method can then be used to compute a three-parameter update, $\phi_k$ via the Moore-Penrose psuedoinverse:
    \begin{equation}
    	\phi_k = ({\nabla r}^T {\nabla r})^{-1} {\nabla r}^T r(q_k).
    \end{equation}
	The update is then applied using the composition for the group:
	\begin{equation}
		\q_{k+1} = \q_k \otimes \varphi(\phi_k)
	\end{equation}.
	This ``multiplicative'' Gauss-Newton method is summarized in Algorithm \ref{alg:mgn}.

@@

\section{Trajectory Optimization on $\R^n \times SO(3)$} \label{sec:trajopt}
    Here we outline the modifications to the ALTRO solver \cite{howell2019altro} to
    solve trajectory optimization problems for rigid bodies, which extends easily to
    arbitrary systems whose state is in $\R^n \times SO(3)$. %ALTRO is an efficient solver for constrained nonlinear optimization problems that uses iterative LQR (iLQR) with an augmented Lagrangian framework.

    We consider trajectory optimization problems of the form,
    \begin{mini}[2]
        {x_{0:N},u_{0:N-1}}{\ell_f(x_N) + \sum_{k=0}^{N-1} \ell_k(x_k,u_k) }{}{}
        \addConstraint{x_{k+1} = f(x_k,u_k)}
        \addConstraint{g_k(x_k,u_k)}{\leq 0}
        \addConstraint{h_k(x_k,u_k)}{=0}
        \label{discrete_trajopt},
    \end{mini}
    where $x$ and $u$ are the state and control vectors as described in Sec. \ref{sec:rigidbody_dynamics},
    $f$ are the dynamics as defined in \eqref{eq:rigid_body_dynamics}, $\ell_k$ is a general
    nonlinear cost function at a single time step,
    $N$ is the number of time steps, and $g_k$, $h_k$ are general nonlinear inequality and 
    equality constraints.

    Like most gradient or Newton-based methods for optimization, ALTRO approximates the
    nonlinear functions $f, \ell, g,$ and $h$ with their first or second-order Taylor
    series expansions. Leveraging the methods from Sec: \ref{sec:Quaternion_Calculus}, we
    adapt the algorithm to optimize directly on the error state $\delta x \in \R^{12}$:
    \begin{equation} \label{eq:state_error}
    	\delta x_k = \begin{bmatrix} 
            r_k - \bar{r}_k \\ \varphi^{-1}(\bar{\q}_k^{-1} \otimes \q_k) \\ v_k - \bar{v}_k \\ \omega_k - \bar{\omega}_k 
        \end{bmatrix} .
    \end{equation}

@@

	With these results, we can apply standard Newton and quasi-Newton techniques along the lines of Section \ref{sec:Wahbas}. We can also calculate a second-order expansion of the ``action-value function'' $Q(x,u)$ needed in differential dynamic programming (DDP) and LQR-based methods,
    \begin{align}
        Q_{xx} &= \ell_{xx} + A_{k}^T P_{k+1} A_{k} \label{Qxx_exp}\\
        Q_{uu} &= \ell_{uu} + B_{k}^T P_{k+1} B_{k} \label{Quu_exp}\\
        Q_{ux} &= \ell_{ux} + B_{k}^T P_{k+1} A_{k} \label{Qux_exp}\\
        Q_x &= \ell_x  + A_{k}^T p_{k+1} \label{Qx_exp}\\
        Q_u &= \ell_u + B_{k}^T p_{k+1} \label{Qu_exp},
    \end{align}
    from which we can calculate the quadratic expansion of the cost-to-go 
    $P_k \in \R^{12 \times 12}$, $p_k \in \R^{12}$, and optimal linear feedback gains 
    $K_k \in \R^{m \times 12}$ and feed-forward corrections $d_k \in \R^m$ by starting at the terminal state and
    performing a backward Riccati recursion as usual~\cite{li2004iterative,howell2019altro}.

    During the ``forward rollout'' of these methods, the dynamics are simulated forward in time using updated control inputs: 
    \begin{equation} \label{eq:mlqr_control}
        u_k = \bar{u}_k - d_k - K_k \delta x_k.
    \end{equation}
    where $\bar{u}_k$ is the control value from the previous iteration, and $\delta x$ is
    computed using \eqref{eq:state_error}, with $x_k$ being the current state estimate
    and $\bar{x}_k$ the state from the previous iteration. 
    For more details on the ALTRO algorithm, which leverages ideas from both DDP and
    direct Newton methods, we refer the reader to~\cite{howell2019altro}.

@@

    \subsection{Quadrotor Flip}
        A 360 degree flip trajectory for a quadrotor was optimized with dynamics adapted
        from \cite{mellinger2012trajectory}. To encourage the flip, we specified a 
        ``waypoint'' cost function of the following form:
        \begin{multline}
                  \sum_{k \in \mathcal{N}} \ell(x_k, \hat{x}, \hat{Q}) 
                + \sum_{k \in \mathcal{W}} \ell(x_k, \bar{x}_k, Q_w) 
                + \half \sum_{k=1} u_k^T R u_k
        \end{multline}
        with $\mathcal{W} = \{45,51,55,75,101\}$, $\mathcal{N} = \{1:101\} \setminus
        \mathcal{W}$. The four intermediary ``waypoints'' were used to encourage the
        quadrotor to reach angles of \ang{90}, \ang{180}, \ang{270}, and \ang{360} around
        an approximately circular arc. The last waypoint was used to encourage the
        quadrotor to move towards the final goal. The quadrotor was constrained to stay
        above the floor and move to a goal state 2 meters away in the $+y$ direction. The
        solver was provided a dynamically infeasible initial trajectory that linearly
        interpolates between the initial and final states, rotating the quadrotor around
        the x-axis a full \ang{360}.

@@

\section*{Acknowledgements}
This material is based upon work supported by the National Science Foundation Graduate
Research Fellowship Program under Grant No. DGE-1656518. Any opinions, findings, and
conclusions or recommendations expressed in this material are those of the author(s) and
do not necessarily reflect the views of the National Science Foundation

\todo{Add NASA acknowledgements (JPL + ECF)}

>>>>

    The Jacobian of $r(q)$ can be found using \eqref{eq:quat_jacobian} and \eqref{eq:quaternion_rotation}: \todo{Kevin: fix the second line of this equation:}
    \begin{equation} \label{eq:newton_foc}
        \begin{aligned}
            \nabla r(q) &= \pdv{r}{q} G(q) \\
              &= \left( \sum_i -2 H^T R(q)^T R({}^B \hat{v}_i) \right).
        \end{aligned}
    \end{equation}
    Given a guess solution, $q_k$, The standard Gauss-Newton method can then be used to compute a three-parameter update $\phi_k$ via the Moore-Penrose psuedoinverse:
    \begin{equation}
    	\phi_k = ({\nabla r}^T {\nabla r})^{-1} {\nabla r}^T r(q_k).
    \end{equation}
	The update is then applied using the composition for the group:
	\begin{equation}
		\q_{k+1} = \q_k \otimes \varphi(\phi_k)
	\end{equation}.
	This ``multiplicative'' Gauss-Newton method is summarized in Algorithm \ref{alg:mgn}.

@@

\section{Trajectory Optimization on $\R^n \times SO(3)$} \label{sec:trajopt}
    Here we outline the modifications to the ALTRO solver \cite{howell2019altro} to
    solve trajectory optimization problems for rigid bodies, which extends easily to
    arbitrary systems whose state is in $\R^n \times SO(3)$. %ALTRO is an efficient solver for constrained nonlinear optimization problems that uses iterative LQR (iLQR) with an augmented Lagrangian framework.
    We consider trajectory optimization problems of the form,
    \begin{mini}[2]
        {x_{0:N},u_{0:N-1}}{\ell_f(x_N) + \sum_{k=0}^{N-1} \ell_k(x_k,u_k) }{}{}
        \addConstraint{x_{k+1} = f(x_k,u_k)}
        \addConstraint{g_k(x_k,u_k)}{\leq 0}
        \addConstraint{h_k(x_k,u_k)}{=0}
        \label{discrete_trajopt},
    \end{mini}
    where $x$ and $u$ are the state and control vectors as described in Sec. \ref{sec:rigidbody_dynamics},
    $f$ are the dynamics as defined in \eqref{eq:rigid_body_dynamics}, $\ell_k$ is a general
    nonlinear cost function at a single time step,
    $N$ is the number of time steps, and $g_k$, $h_k$ are general nonlinear inequality and 
    equality constraints.

    ALTRO combines techniques from both differential dynamic programming (DDP) and direct transcription methods to achieve high performance on challenging constrained nonlinear trajectory optimization problems. Like most methods for nonlinear optimization, ALTRO iteratively approximates the
    nonlinear functions $f, \ell, g,$ and $h$ with their first or second-order Taylor
    series expansions. Leveraging the methods from Sec: \ref{sec:Quaternion_Calculus}, we
    adapt the algorithm to optimize directly on the error state $\delta x \in \R^{12}$:
    \begin{equation} \label{eq:state_error}
    	\delta x_k = \begin{bmatrix} 
            r_k - \bar{r}_k \\ \varphi^{-1}(\bar{\q}_k^{-1} \otimes \q_k) \\ v_k - \bar{v}_k \\ \omega_k - \bar{\omega}_k 
        \end{bmatrix} .
    \end{equation}

@@

	With these results, we can apply standard Newton and quasi-Newton techniques along the lines of Section \ref{sec:Wahbas}. We can also calculate a second-order expansion of the ``action-value function'' $Q(x,u)$ needed in DDP and LQR-based methods,
    \begin{align}
        Q_{xx} &= \ell_{xx} + A_{k}^T P_{k+1} A_{k} \label{Qxx_exp}\\
        Q_{uu} &= \ell_{uu} + B_{k}^T P_{k+1} B_{k} \label{Quu_exp}\\
        Q_{ux} &= \ell_{ux} + B_{k}^T P_{k+1} A_{k} \label{Qux_exp}\\
        Q_x &= \ell_x  + A_{k}^T p_{k+1} \label{Qx_exp}\\
        Q_u &= \ell_u + B_{k}^T p_{k+1} \label{Qu_exp},
    \end{align}
    from which we can calculate the quadratic expansion of the cost-to-go 
    $P_k \in \R^{12 \times 12}$, $p_k \in \R^{12}$, and optimal linear feedback gains 
    $K_k \in \R^{m \times 12}$ and feed-forward corrections $d_k \in \R^m$ by starting at the terminal state and
    performing a backward Riccati recursion as usual~\cite{li2004iterative,howell2019altro}.

    During the ``forward rollout'' of these methods, the dynamics are simulated forward in time using updated control inputs: 
    \begin{equation} \label{eq:mlqr_control}
        u_k = \bar{u}_k - d_k - K_k \delta x_k.
    \end{equation}
    where $\bar{u}_k$ is the control value from the previous iteration, and $\delta x$ is
    computed using \eqref{eq:state_error}.%, with $x_k$ being the current state estimate and $\bar{x}_k$ the state from the previous iteration. 
    For more details on the ALTRO algorithm, we refer the reader to~\cite{howell2019altro}.

@@

    \subsection{Quadrotor Flip}
        A 360 degree flip trajectory for a quadrotor was optimized with dynamics adapted
        from \cite{mellinger2012trajectory}. To encourage the flip, we specified a 
        ``waypoint'' cost function of the following form:
        \begin{multline}
                  \sum_{k \in \mathcal{N}} \ell(x_k, \hat{x}, \hat{Q}) 
                + \sum_{k \in \mathcal{W}} \ell(x_k, \bar{x}_k, Q_w) 
                + \half \sum_{k=1} u_k^T R u_k
        \end{multline}
        with $\mathcal{W} = \{45,51,55,75,101\}$, $\mathcal{N} = \{1:101\} \setminus
        \mathcal{W}$. The four intermediary ``waypoints'' were used to encourage the
        quadrotor to reach angles of \ang{90}, \ang{180}, \ang{270}, and \ang{360} around
        an approximately circular arc. The last waypoint was used to encourage the
        quadrotor to move towards the final goal. The quadrotor was constrained to stay
        above the floor and move to a goal state two meters away in the $+y$ direction. The
        solver was provided a dynamically infeasible initial trajectory that linearly
        interpolates between the initial and final states, rotating the quadrotor around
        the x-axis a full \ang{360}.

@@

\section*{Acknowledgements}
This work was supported by a NASA Early Career Faculty Award (Grant Number 80NSSC18K1503). This material is based upon work supported by the National Science Foundation Graduate
Research Fellowship Program under Grant No. DGE-1656518. Any opinions, findings, and
conclusions or recommendations expressed in this material are those of the author(s) and
do not necessarily reflect the views of the National Science Foundation.
